---
title: "Cannabis Monthly Permitting Report"
output:
  word_document:
    reference_docx: report_template.docx
  html_document:
    df_print: paged
  pdf_document: default
params:
  prog_date:
    label: Beginning of program
    input: date
    value: "2018-01-01"
  year_date:
    label: Beginning of year date
    input: date
    value: "2021-01-01"
  start_date:
    label: Begining of month date
    input: date
    value: "2021-11-01"
  end_date:
    label: End of month date
    input: date
    value: "2021-11-30"
---

```{r global_options, include = FALSE}

knitr::opts_chunk$set(echo = TRUE, fig.show = 'asis')

```


```{r, include = FALSE}

# Clear workspace -- commented out for knitting
#rm(list = ls())

# Load required packages and scripts --------------------------- 

library(tidyverse)
library(lubridate)
library(rlang)
library(knitr)
library(gdtools)
library(flextable)
library(officer)
library(tcltk)
library(plotly)
source("../LSA_Functions.R")

```


```{r, include = FALSE}

# Set date parameters --------------------------- 

prog.date = params$prog_date
year.date = params$year_date
start.date = params$start_date
end.date = params$end_date

#defines a date stamp format to convert numerical format dates
ds <- stamp("January 1, 2022")


# Create date labels --------------------------- 

mo.label = str_c(as.character(month(end.date, label = TRUE, abbr = FALSE)),
                 as.character(year(end.date)), sep = " ")

ytd.label = str_c(as.character(year(end.date)), "Year-to-Date", sep = " ")

```


```{r, include = FALSE}

# User choose files --------------------------- 

# Another option is rstudioapi::selectFile but doesn't currently support multiple file selection

eom_csv = tk_choose.files(caption = "Select end of the month csv file", multi = FALSE, filters = matrix(c("Comma Delimited Files (.csv)", "*.csv"), 1, 2))

```


```{r, include = FALSE}

# Read in input csvs --------------------------- 

eom.in = read_csv(eom_csv, col_names = TRUE, col_types = cols(.default = "c"))

```


```{r, include = FALSE}

# drop unnecessary columns
# remove duplicate rows
# convert dates
# drop rows with status of agreement suspended or revoked
eom.df = eom.in %>%
  select(`ID Number`, Region, Opportunity, `Date Submitted`, Status, `Date of Status`, `Date Issued`, County) %>%
  distinct(`ID Number`, .keep_all = TRUE) %>%
  mutate(across(.cols = contains("Date"), ~as_date(., format = "%m/%d/%Y"))) %>%
  filter(!Status %in% c('Agreement Revoked', 'Agreement Suspended'))

```


```{r, include = FALSE}

# Add opportunity abbrev --------------------------- 

eom.df = eom.df %>%
  mutate(Opp = case_when(
    Opportunity == "Cannabis Cultivation LSA Notification" ~ "LSA-Notify",
    Opportunity == "Apply for Self-Certification (Cannabis Only)" ~ "Self-Cert",
    Opportunity == "Notify for General Agreement (Cannabis Only)" ~ "GA",
    Opportunity == "Standard Agreement" ~ "SA",
    str_detect(Opportunity, "Standard Agreement") ~ "SA",
    str_detect(Opportunity, "Paper Notification") ~ "SA"
  ))

```


```{r, include = FALSE}

# Create sub-category groupings --------------------------- 

df = define_category(eom.df)

# Change notifications under review date to end of month report date
# Some permits have status date that are months old, so need date to reflect currently under review
# note: use if_else to preserve date class
df = df %>%
  mutate(Date = if_else(Metric == "Notifications in Progress", ymd(end.date), Date),
         Year = if_else(Metric != "Notifications in Progress", Year, as.factor(year(end.date))))

```


```{r, include = FALSE, echo = FALSE, results = "asis"}

# Get counts ---------------------------

# Filter to month
mo = df %>% 
  filter(Date >= start.date & Date <= end.date) %>%  
  mutate(DateRange = mo.label)

# Filter to year
ytd.filter = str_c(as.character(year(end.date)), "-01-01")
ytd = df %>% 
  filter(Date >= ytd.filter) %>%
  mutate(DateRange = ytd.label)

# Program to date
ptd = df %>% 
  mutate(DateRange = "Program-to-Date")

# Combine all date categories
df.g = bind_rows(mo, ytd, ptd)

# Counts by region
n.reg = df.g %>%
  group_by(Metric, Category, Region, DateRange) %>%
  summarise(n = n(), .groups = "keep") %>% 
  ungroup() %>% 
  complete(Region, DateRange, nesting(Metric, Category), fill = list(n = 0)) %>%
  filter(!((DateRange == ytd.label | DateRange == "Program-to-Date") & Metric %in% c("Notifications in Progress", "Draft Agreement Issued"))) %>% 
  mutate(Group = "Region")

# Counts by county
n.co = df.g %>%
  group_by(Metric, Category, County, DateRange) %>%
  summarise(n = n(), .groups = "keep") %>% 
  ungroup() %>% 
  complete(County, DateRange, nesting(Metric, Category), fill = list(n = 0)) %>%
  filter(!((DateRange == ytd.label | DateRange == "Program-to-Date") & Metric %in% c("Notifications in Progress", "Draft Agreement Issued"))) %>% 
  mutate(Group = "County")

# Counts for entire program (i.e., not grouped by county or region)
n.all = df.g %>%
  group_by(Metric, Category, DateRange) %>%
  summarise(n = n(), .groups = "keep") %>% 
  ungroup() %>% 
  complete(DateRange, nesting(Metric, Category), fill = list(n = 0)) %>%
  filter(!((DateRange == ytd.label | DateRange == "Program-to-Date") & Metric %in% c("Notifications in Progress", "Draft Agreement Issued"))) %>% 
  mutate(Group = "Program-Wide")

# Counts by region for month (include in progress sub-metrics)
n.reg.mo = mo %>%
  group_by(Metric, SubMetric, Category, Region, DateRange) %>%
  summarise(n = n(), .groups = "keep") %>% 
  ungroup() %>% 
  complete(Region, DateRange, nesting(Metric, SubMetric, Category), fill = list(n = 0)) %>%
  filter(!((DateRange == ytd.label | DateRange == "Program-to-Date") & Metric %in% c("Notifications in Progress", "Draft Agreement Issued"))) %>% 
  mutate(Group = "Region")


```


```{r, include = FALSE, results = "asis"}

# Combine count data and save to csv ---------------------------

out.csv = str_c("CannabisMonthlyReport", "_", as.character(month(end.date, label = TRUE, abbr = TRUE)), as.character(year(end.date)), ".csv", sep = "")

bind_rows(n.reg, n.co, n.all) %>%
  select(Category, Metric, DateRange, Group, Region, County, n) %>%
  rename(Count = n) %>%
  write_csv(out.csv, col_names = TRUE)

```


```{r, include = FALSE}

# Create plots ---------------------------

# Plot data

plot.data = n.reg %>% 
  group_by(Region, DateRange, Metric) %>% 
  summarise(n = sum(n, na.rm = TRUE), .groups = "keep") %>%
  group_by(DateRange, Metric) %>%
  mutate(prop = n / sum(n)) %>% 
  arrange(desc(Region)) %>%
  #mutate(lab.ypos = cumsum(prop) - (0.5 * prop)) %>%
  ungroup()

plot.data.mo = plot.data %>% 
  filter(DateRange == mo.label, Metric %in% c('Notifications Received', 'Notifications in Progress', 'Final Document Issued'))
plot.data.ytd = plot.data %>% filter(DateRange == ytd.label) %>% 
  filter(DateRange == ytd.label, Metric %in% c('Notifications Received', 'Final Document Issued'))
plot.data.to.date = plot.data %>% filter(DateRange == "Program-to-Date") %>% 
  filter(DateRange == 'Program-to-Date', Metric %in% c('Notifications Received', 'Final Document Issued'))


# Plot function

# reg.colors = c("#D53E4F", "#F46D43", "#FDAE61", "#ABDDA4", "#66C2A5", "#3288BD")
# https://github.com/wilkox/ggfittext
# https://ggrepel.slowkow.com/articles/examples.html
# https://stackoverflow.com/questions/36561030/how-expand-ggplot-bar-scale-on-one-side-but-not-the-other-without-manual-limits

make_region_barplot = function(in.df, suffix){
  
  in.df = in.df %>% 
    mutate(Metric = fct_relabel(Metric, ~ paste(.x, suffix, sep = ' ')),
           Metric = fct_drop(Metric))
  
  p = ggplot(data = in.df, aes(x = Region, y = n, fill = Region)) +
    geom_col() +
    #scale_fill_brewer(palette = 'Set2') +
    scale_fill_viridis_d(alpha = 0.8) +
    #geom_text(aes(label = scales::percent(pct)) +
    geom_text(aes(label = paste0(round(prop * 100), '%')),
              colour = 'grey30', hjust = -0.15, fontface = 'italic', size = 3) +
    scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
    labs(y = 'Count', x = NULL) +
    coord_flip() +
    facet_wrap(vars(Metric), ncol = 1) +
    theme_minimal() +
    theme(legend.position = "none",
          axis.title.y = element_blank(),
          axis.title.x = element_text(color = 'grey30'),
          axis.text = element_text(color = 'grey30'),
          panel.grid.major.y = element_blank(),
          strip.background = element_blank(),
          strip.text.x = element_text(hjust = 0, face = 'plain', color = 'grey30', size = rel(1.5)))
  
  return(p)
  
}

# Generate plots

plot.mo = make_region_barplot(plot.data.mo, '')
plot.ytd = make_region_barplot(plot.data.ytd, 'YTD')
plot.to.date = make_region_barplot(plot.data.to.date, 'PTD')

```


# `r mo.label`

## `r ds(ymd(start.date))` to `r ds(ymd(end.date))`


```{r, include = FALSE, echo = FALSE, results = "asis"}

# This month by region ---------------------------

mo.reg = n.reg.mo %>%
  filter(DateRange == mo.label) %>%
  pivot_wider(names_from = Region, values_from = n) %>%
  select(Metric, SubMetric, Category, R1, R2, R3, R4, R5, R6) %>%
  # rows sums to new column
  mutate(Total = rowSums(across(R1:R6))) %>% 
  # column sum to new rows
  group_by(Metric, SubMetric) %>% 
  bind_rows(summarise(.,
                      across(where(is.numeric), sum),
                      across(where(is.factor), ~"Total"),
                      .groups = "keep")) %>%
  ungroup() 

# add notifications in progress totals
prog.total = mo.reg %>%
  filter(Metric == "Notifications in Progress", Category == "Total") %>%
  group_by(Metric) %>%
  summarise(across(R1:R6, sum), .groups = "keep") %>%
  ungroup() %>%
  mutate(Total = rowSums(across(R1:R6)),
         Category = "Total") 

submet.ord <- c("Incomplete", "Under Review", "Complete", "")

cat.ord <- c("SA", "Self-Cert", "GA", "SA Issued", "Self-Cert Letter", "GA Issued", "OpLaw", "Notification Not Required", "No Agreement Needed", "", "Total")

mo.reg = mo.reg %>%
  bind_rows(prog.total) %>%
  mutate(Category = factor(Category, levels=cat.ord),
         SubMetric = fct_explicit_na(SubMetric, ""),
         SubMetric = fct_relevel(SubMetric, "", after = Inf)) %>%
  arrange(Metric, SubMetric, Category)


# fixes the border issue by changing cells to two spaces to differentiate
mo.reg$SubMetric <- as.character(mo.reg$SubMetric)
mo.reg$SubMetric[mo.reg$Metric ==  "Draft Agreement Issued"] <- "  "
mo.reg$SubMetric <- as.factor(mo.reg$SubMetric)

mo.reg = subset(mo.reg,!(Metric == "Draft Agreement Issued" & Category == "  "))

```


```{r, include = TRUE, echo = FALSE, results = "asis"}

# Table for this month by region ---------------------------


mo.reg %>%
  flextable() %>%
  merge_v(j = "Metric") %>%
  merge_v(j = ~SubMetric) %>%
  valign(j = "Metric", valign = "top") %>%
  valign(j = "SubMetric", valign = "top") %>%
  font(fontname = "Source Sans Pro Light", part = "all") %>%
  fontsize(size = 11, part = "all") %>%
  #bold and italicize row totals
  bold(j = 3:ncol(mo.reg), i = which(mo.reg$Category == "Total")) %>% 
  italic(j = 3:ncol(mo.reg), i = which(mo.reg$Category == "Total")) %>%
  #bold last column (total)
  bold(j = "Total", part = "all") %>%
  italic(j = "Total", part = "all")%>%
  bold(j = "Metric", part = "all") %>%
  autofit() %>%
  width(., width = dim_pretty(.)$widths) %>%
  #alternating row color, but makes columns 1 and 2 all grey
  theme_zebra() %>% 
  bg(j = 1:2, part = "body", bg = "#EFEFEF") %>% 
  border_outer(border = fp_border(width=2)) %>% 
  border_inner_h() %>%
  border_inner_v() %>%
  hline(i = which(mo.reg$Category == "Total"), border = fp_border(width=2)) %>%
  hline_bottom(part = "header", border = fp_border(width=2)) %>%
  fix_border_issues()


```



```{r, include = TRUE, echo = FALSE, fig.width = 7.5, fig.height = 6.5, fig.keep = 'high', dpi = 500, eval = TRUE}

plot(plot.mo)

```
Figure 1: Monthly metrics shown as regional percentages of statewide totals.
\pagebreak

# `r ytd.label`

## `r ds(ymd(year.date))`  to `r ds(ymd(end.date))`
 

```{r, include = FALSE}

# YTD by region ---------------------------


ytd.reg = n.reg %>%
  filter(DateRange == ytd.label) %>%
  pivot_wider(names_from = Region, values_from = n) %>%
  select(Metric, Category, R1, R2, R3, R4, R5, R6) %>%
  mutate(Total = rowSums(across(R1:R6))) %>% 
  group_by(Metric) %>% 
  bind_rows(summarise(.,
                      across(where(is.numeric), sum),
                      across(where(is.factor), ~"Total"),
                      .groups = "keep")) %>%
  ungroup() %>%
  mutate(Category = factor(Category, levels=cat.ord)) %>%
  arrange(Metric, Category)

```


```{r, include = TRUE, echo = FALSE, results = "asis"}

# Table for ytd by region ---------------------------

ytd.reg %>% 
  flextable() %>%
  merge_v(j = "Metric") %>%
  valign(j = "Metric", valign = "top") %>%
  font(fontname = "Source Sans Pro Light", part = "all") %>%
  fontsize(size = 11, part = "all") %>%
  #bold and italicize row totals
  bold(j = 2:ncol(ytd.reg), i = which(ytd.reg$Category == "Total")) %>% 
  italic(j = 2:ncol(ytd.reg), i = which(ytd.reg$Category == "Total",)) %>% 
  bold(j = "Total", part = "all") %>%
  italic(j = "Total", part = "all")%>%
  bold(j = "Metric", part = "all") %>%
  autofit() %>%
  width(., width = dim_pretty(.)$widths) %>%
  theme_zebra() %>% 
  #bg(j = 1, part = "body", bg = "#EFEFEF") %>% 
  border_outer(border = fp_border(width=2)) %>% 
  border_inner_h() %>%
  border_inner_v() %>%
  hline(i = which(ytd.reg$Category == "Total"), border = fp_border(width=2)) %>%
  hline_bottom(part = "header", border = officer::fp_border(width=2)) %>%
  fix_border_issues()

```

```{r, include = TRUE, echo = FALSE, fig.width = 7.5, fig.keep = 'high', dpi = 500, eval = TRUE}

plot(plot.ytd)

```
Figure 2: Year-to-Date metrics shown as regional percentages of statewide totals.
\pagebreak

# Program-to-Date

## `r ds(ymd(prog.date))`  to `r ds(ymd(end.date))`


```{r, include = FALSE}

# Program count by region--------------------------- 

ptd.reg = n.reg %>%
  filter(DateRange == "Program-to-Date") %>%
  pivot_wider(names_from = Region, values_from = n) %>%
  select(Metric, Category, R1, R2, R3, R4, R5, R6) %>%
  mutate(Total = rowSums(across(R1:R6))) %>% 
  group_by(Metric) %>% 
  bind_rows(summarise(.,
                      across(where(is.numeric), sum),
                      across(where(is.factor), ~"Total"),
                      .groups = "keep")) %>%
  ungroup() %>%
  mutate(Category = factor(Category, levels=cat.ord)) %>%
  arrange(Metric, Category)

```


```{r, include = TRUE, echo = FALSE, results = "asis"}

# Table for program to date ---------------------------

ptd.reg %>% 
  flextable() %>%
  merge_v(j = "Metric") %>%
  valign(j = "Metric", valign = "top") %>%
  font(fontname = "Source Sans Pro Light", part = "all") %>%
  fontsize(size = 11, part = "all") %>%
  bold(j = 2:ncol(ptd.reg), i = which(ptd.reg$Category == "Total")) %>% 
  italic(j = 2:ncol(ptd.reg), i = which(ptd.reg$Category == "Total",)) %>% 
  bold(j = "Total", part = "all") %>%
  italic(j = "Total", part = "all")%>%
  bold(j = "Metric", part = "all") %>%
  autofit() %>%
  width(., width = dim_pretty(.)$widths) %>%
  theme_zebra() %>% 
  #bg(j = 1, part = "body", bg = "#EFEFEF") %>% 
  border_outer(border = fp_border(width=2)) %>% 
  border_inner_h() %>%
  border_inner_v() %>%
  hline(i = which(ptd.reg$Category == "Total"), border = fp_border(width=2)) %>%
  hline_bottom(part = "header", border = fp_border(width=2)) %>%
  fix_border_issues()

```


```{r, include = TRUE, echo = FALSE, fig.width = 7.5, fig.keep = 'high', dpi = 500, eval = TRUE}

plot(plot.to.date)

```
Figure 3: Program-to-Date metrics shown as regional percentages of statewide totals.


```{r, include = TRUE, echo = FALSE, fig.keep = 'high', eval = FALSE}

#https://stackoverflow.com/questions/37285729/how-to-give-subtitles-for-subplot-in-plot-ly-using-r
# https://stackoverflow.com/questions/64108716/bold-plotly-r-pie-chart-labels
#https://plotly-r.com/exporting-static-images.html
#https://stackoverflow.com/questions/52258420/unexpected-behaviour-in-ggplot2-pie-chart-labeling
#https://stackoverflow.com/questions/47752037/pie-chart-with-ggplot2-with-specific-order-and-percentage-annotations/47752576
#https://ggrepel.slowkow.com/articles/examples.html

fig = plot_ly()
fig = fig %>%
  add_pie(data = plot.data.ytd %>% filter(Metric == "Notifications Received"), 
          labels = ~Region, values = ~n,
          hole = 0.6,
          texttemplate="%{label}<br>%{percent:1%}",
          textfont = list(color = '#000000'),
          #textposition = "outside",
          insidetextorientation = "horizontal",
          sort = FALSE,
          title = 'Notifications Received', 
          marker = list(colors = reg.colors,
                       line = list(color = '#FFFFFF', width = 2)),
          domain = list(row = 0, column = 0))
fig = fig %>%
   add_pie(data = plot.data.ytd %>% filter(Metric == "Final Document Issued"), 
          labels = ~Region, values = ~n,
          hole = 0.6,
          title = 'Final Document Issued',
          texttemplate="%{label}<br>%{percent:1%}",
          insidetextorientation = "horizontal",
          sort = FALSE,
          marker = list(colors = reg.colors,
                       line = list(color = '#FFFFFF', width = 2)),
          domain = list(row = 0, column = 1))
fig = fig %>%
  layout(showlegend = FALSE,
         grid = list(rows = 1, columns = 3),
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         width = 900, height = 500,
         template = 'plotly_dark')

fig

```

