---
title: "DCC_LSA_UID_Cleanup"
author: "Caroline Larsen"
date: "2022-08-26"
output:
  word_document: default
  html_document:
    theme: yeti
    highlight: kate
    df_print: paged
    code_folding: hide
  pdf_document: default
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = TRUE)
```

### Workspace set-up

1.  Clean current work space
2.  Suppress scientific notation
3.  Load required packages

```{r echo=FALSE}

# Clean current work space
rm(list=ls())

# Suppress scientific notation
options(scipen = 9999)

# Load required package
library(tidyverse)
library(lubridate)
library(dbplyr)
library(tcltk)
library(rlang)
library(knitr)
library(stringr)
library(tinytex)

```

### User selects input file

1.  User selects input file of most recent DCC licenses with LSA UIDs
    that need cleaning. The code should work on any CSV file so long as
    the original LSA UID column name is "LSA ID Number".
2.  Upload input file of opportunity ID names from SharePoint located in
    SPECIFYLOCATIONHERE

```{r}
# Choose file path for recent DCC license information
dcc.csv = tk_choose.files(caption = "Please select recent DCC LICENSE DATA CSV FILE")

# Choose file path for opportunity ID information
opp.id.csv = tk_choose.files(caption = "Please select OPPORTUNITY ID CSV FILE")


# Read in the files
dcc.in <- read_csv(dcc.csv, col_types = cols(.default = "c"), na = c("", "NA", "NULL"))
head(as.data.frame(dcc.in))
dcc <- dcc.in

opp.id.in <- read_csv(opp.id.csv, col_types = cols(.default = "c"), na = c("", "NA", "NULL"))
head(as.data.frame(opp.id.in))
opp.id <- opp.id.in
opp.lst <- opp.id$ID


```

### 1a. Definite acceptable formats and check proportions correct/incorrect

List of acceptable LSA UID formats 1. e.g. EPIMS-ABC-12345-R2 2. e.g.
EPIMS-ABC-12345-R1C 3. e.g. EPIMS-12345-R2 4. e.g. EPIMS-12345-R1C 5.
e.g. 12345 6. e.g. ABC-12345 7. e.g. 1600-2000-0000-R2 8. e.g.
1600-2000-0000-R1C

```{r}

accpt.id <- c("^EPIMS-[:alpha:]{3}-[:digit:]{5}-R\\d+$", 
              "^EPIMS-[:alpha:]{3}-[:digit:]{5}-R\\d+[:alpha:]+$",
              "^EPIMS-[:digit:]{5}-R\\d+$",
              "^EPIMS-[:digit:]{5}-R\\d+[:alpha:]+$",
              "^[:digit:]{5}$",
              "^[:alpha:]{3}-[:digit:]{5}$",
              "^1600-2[:digit:]{3}-[:digit:]{4}-0000-R\\d+$",
              "^1600-2[:digit:]{3}-[:digit:]{4}-0000-R\\d+[:alpha:]+$")

```

### 1b. Function to count "correct" vs "incorrect" LSA UIDs and print "incorrect"

```{r}

incorrect_lsa = function(dat, columnname) {
  
  accpt.id <- c("^EPIMS-[:alpha:]{3}-[:digit:]{5}-R\\d{1}$",
                "^EPIMS-[:alpha:]{3}-[:digit:]{5}-R\\d{1}[:alpha:]{1}$$",
                "^EPIMS-[:digit:]{5}-R\\d{1}$",                             
                "^EPIMS-[:digit:]{5}-R\\d{1}[:alpha:]{1}$",
                "^[:digit:]{5}$",
                "^[:alpha:]{3}-[:digit:]{5}$",
                "^1600-2[:digit:]{3}-[:digit:]{4}-0000-R\\d{1}$",
                "^1600-2[:digit:]{3}-[:digit:]{4}-0000-R\\d{1}[:alpha:]{1}$")
  
  dat.filt <- dat %>%
    filter(str_detect(columnname, paste(accpt.id, collapse = "|")))
  
  dat.filtrev <- dat %>%
    filter(!str_detect(columnname, paste(accpt.id, collapse = "|")))
  
  print(paste("total: ", nrow(dat)))
  print(paste("correct: ", length(dat.filt$`LSA ID Number`)))
  print(paste("incorrect: ",length(dat.filtrev$`LSA ID Number`)))
  
  if (nrow(dat.filtrev) <= 500) {print(dat.filtrev$`LSA ID Number`)}
  
}

# Check correction status
incorrect_lsa(dcc, dcc$`LSA ID Number`)

```

### 1c. Create a new column for cleaned UIDs, initially populated with submitted LSA UID numbers

```{r}

dcc$LSA.corrected <- dcc$`LSA ID Number`

```

### 2. Note LSA UIDs that contain Opportunity UIDs

Identifies LSA UIDs that match opportunity IDs and marks them as "NA -
Opportunity ID used" in the new LSA.corrected column

```{r}

dcc$LSA.corrected <- lapply(dcc$LSA.corrected, function (x) replace(x, grep(paste(".*", paste(opp.lst, collapse = "|"), ".*"), x), "NA - Opportunity ID used"))

#Brief in-R report
LSA.Opp.Used <- as.data.frame(subset(dcc, LSA.corrected == "NA - Opportunity ID used"))
LSA.Opp.Used <- cbind(LSA.Opp.Used$`LSA ID Number`, unlist(LSA.Opp.Used$LSA.corrected))
LSA.Opp.report <- list()
LSA.Opp.report[[1]] <- "LSA UIDs using Opportunity ID Codes"
LSA.Opp.report[[2]] <- LSA.Opp.Used
print(LSA.Opp.report)

```

### 3. General clean up

This list can be added to easily if common incorrect patterns emerge.
The code runs the following corrections: 1. changes empty space to - 2.
changes \_ . , ; / to - 3. changes -- to - 4. removes - hanging at end
of UIDs 5. removes "(Rev2)" tags at the end of UIDs 6. removes anything
after RX/RXX pattern 7. corrects misspellings of EPIMS 8. adds dash
between "EPIMS" and 5 digit code if missing 9. adds dash between 5 digit
code and region code if missing 10. deletes year from EPIMS IDs 11.
deletes duplicated region code 12. fixes specific typo case where dash
splits 5 digit code into 4 digit code and an extra digit before region
code 13. adds 0000 into 1600 UIDs where missing 14. adds 0000 into 1600
UIDs where missing 15. deletes "EPIMS" from 1600 IDs 16. adds "EPIMS" to
front of IFS with county code + 5 digit code + regional code 17. adds
dashes to EPIMS format UIDs without any (e.g. EPIMS12345R1) 18. drops
"EPIMS" from UIDs formatted e.g. EPIMS-12345 19. deletes three letter
county code from 1600 format UIDs

```{r}

# Correction code
dcc <- dcc %>%  
  mutate(LSA.corrected = str_replace_all(LSA.corrected, " ", "-")) %>%                #1
  mutate(LSA.corrected = str_replace_all(LSA.corrected, ("[_.,;/]"), "-")) %>%        #2
  mutate(LSA.corrected = str_replace_all(LSA.corrected, "--", "-")) %>%               #3
  mutate(LSA.corrected = str_replace_all(LSA.corrected, "(-$)", "")) %>%              #4
  mutate(LSA.corrected = str_replace_all(LSA.corrected, "\\([^)]*\\)", "")) %>%       #5 
  mutate(LSA.corrected = str_replace_all(LSA.corrected, "(.R[:alnum:]{1,2})(-[:alnum:]*$)", "\\1")) %>%     #6 
  mutate(LSA.corrected = str_replace_all(LSA.corrected, "EPMIS|EPMS-|EMPIS|Epims|epims|EPIM-|^PIMS", "EPIMS")) %>%  #7
  mutate(LSA.corrected = str_replace_all(LSA.corrected, "(EPIMS)([:digit:]{5})", "\\1-\\2")) %>%            #8
  mutate(LSA.corrected = str_replace_all(LSA.corrected, "([:digit:]{5})(R)", "\\1-\\2")) %>%                #9
  mutate(LSA.corrected = str_replace_all(LSA.corrected, "(EPIMS)(-20\\d{2})(-\\d{5})", "\\1\\3")) %>%       #10
  mutate(LSA.corrected = str_replace_all(LSA.corrected, "(EPIMS-\\d{5})(R[:alnum:]{1,2})(R[:alnum:]{1,2})", "\\1\\2")) %>%              #11
  mutate(LSA.corrected = str_replace_all(LSA.corrected, "(EPIMS-\\d{4})(-)([:digit:]{1})(R[:alnum:]{1,2})", "\\1\\3\\2\\4")) %>%        #12
  mutate(LSA.corrected = str_replace_all(LSA.corrected, "(1600-2[:digit:]{3}-[:digit:]{4})(-R\\d{1}$)", "\\1-0000\\2")) %>%             #13
  mutate(LSA.corrected = str_replace_all(LSA.corrected, "(1600-2[:digit:]{3}-[:digit:]{4})(-R\\d{1}[:alpha:]{1}$)", "\\1-0000\\2")) %>% #14
  mutate(LSA.corrected = str_replace_all(LSA.corrected, "^(1600-)(EPIMS-)", "\\1")) %>%                     #15
  mutate(LSA.corrected = str_replace_all(LSA.corrected, "(^[:alpha:]{3}-[:digit:]{5}-R)", "EPIMS-\\1")) %>% #16
  mutate(LSA.corrected = str_replace_all(LSA.corrected, "^(EPIMS)([:digit:]{5})(R[:digit:]{1})$", "\\1-\\2-\\3")) %>% #17
  mutate(LSA.corrected = str_replace_all(LSA.corrected, "^(EPIMS-)([:digit:]{5})$", "\\2")) %>% #18
  mutate(LSA.corrected = str_replace_all(LSA.corrected, "^(1600-2[:digit:]{3}-[:digit:]{4})(-[:alpha:]{3})(-R\\d{1}$)", "\\1\\3")) #19

# Check correction status
incorrect_lsa(dcc, dcc$LSA.corrected)

```

### 4. Final clean + extract 5 digit code

Does a final match test for each \$LSA.corrected entry again acceptable
UID list - If matches, keeps it as is - If doesn't match, but contains 5
digit ID, uses 5 digit ID - Otherwise, uses NA

```{r}

# Run match test code
dcc$LSA.corrected <- ifelse(str_detect(dcc$LSA.corrected, paste(accpt.id, collapse = "|")), dcc$LSA.corrected, 
                                 ifelse( (!str_detect(dcc$LSA.corrected, paste(accpt.id, collapse = "|"))) 
                                         & str_detect(dcc$LSA.corrected,"[:digit:]{5}"), 
                                         str_extract(dcc$LSA.corrected, "[:digit:]{5}"), 
                                         "NA"))

# Check correction status
incorrect_lsa(dcc, dcc$LSA.corrected)

```

### 5. Extract 5 digit code into new column with str_extract

Provides a step to add a tab after code to prevent leading zeros from
being dropped once output CSV file is read into Excel

```{r}

# Extract 5 digit code where possible
dcc$LSA.5digit <- str_extract(dcc$LSA.corrected, "[:digit:]{5}")
head(dcc$LSA.5digit)

# Adds a tab after the 5 digit code 
dcc$LSA.5digit <- paste0(dcc$LSA.5digit, "\t")

dcc <- dcc[,c(1:19,36:37,20:35)]

# An alternate option: adds an apostrophe before 5 digit code - uncomment to use
#dcc$LSA.5digit <- paste0("'",dcc$LSA.5digit)

```

### 6. Write file to CSV

```{r}

# Write CSV file
write.csv(dcc, str_c("DCC_LSA_UIDs_Cleanup_", as.character(today()), ".csv"))

```

### 

```{r}

```
